---
title: |
  | \LARGE Homework assignment #2
author: |
  | \large \rm 2021320322 / Minseo Yoon
date: |
  | \rm \today
output:
  pdf_document:
    extra_dependencies: ["kotex"]
fontsize: 11pt
---


# 1

```{r echo=TRUE, warning=FALSE, message=FALSE}
library(gapminder)
?gapminder
library(dplyr)
```

(a)
```{r echo=TRUE}
gapminder %>%
  group_by(continent) %>%
  summarize(n_distinct(country))
```
(b)
```{r echo=TRUE}
gapminder %>%
  filter(continent == 'Europe', year == 1997) %>%
  arrange(gdpPercap) %>%
  head(n=1)
```
Answer: Albania
```{r echo=TRUE}
gapminder %>%
  filter(continent == 'Europe', year == 2007) %>%
  arrange(gdpPercap) %>%
  head(n=1)
```
Answer: Albania

(c)
```{r echo=TRUE}
gapminder %>%
  filter(year >= 1970 & year < 1980) %>%
  group_by(continent) %>%
  summarize(avg = mean(lifeExp))
```
(d)
```{r echo=TRUE}
gapminder %>%
  mutate(gdp = gdpPercap * pop) %>%
  group_by(country) %>%
  summarize(totalgdp = sum(gdp)) %>%
  arrange(desc(totalgdp)) %>%
  head(n=5)
```
(e)
```{r echo=TRUE}
gapminder %>%
  select(country, lifeExp, year) %>%
  filter(lifeExp >= 82)
```
(f)
```{r echo=TRUE}
gapminder %>%
  filter(continent != 'Europe') %>%
  group_by(continent, year) %>%
  summarize(meanPop = mean(pop)) %>%
  arrange(desc(meanPop))
```
Answer: Asia, 2007

# 2

```{r echo=TRUE, warning=FALSE}
library(nycflights13)
?flights
?planes
?weather
```

(a)
```{r echo=TRUE}
flights %>%
  group_by(month) %>%
  summarize(cancelled = sum(is.na(dep_time)), total = n(), 
            prop = cancelled / total) %>%
  arrange(prop)
```
Highest: February, Lowest: October

It is estimated that flights are often cancelled in summer and winter due to bad weather (rain, snow, and so on).

(b)
```{r echo=TRUE, warning=FALSE, message=FALSE}
library(ggplot2)
```
```{r echo=TRUE}
flights %>%
  filter(year == 2013 & !is.na(tailnum)) %>%
  group_by(tailnum) %>%
  summarize(num = n()) %>%
  arrange(desc(num)) %>%
  head(n=1)
```
```{r echo=TRUE, warning=FALSE, massage=FALSE}
library(lubridate)
```
```{r echo=TRUE}
flights %>%
  filter(tailnum == 'N725MQ') %>%
  mutate(date = paste(sprintf('%04d-%02d-%02d', year, month, day)),
         weeks = week(date)) %>%
  group_by(weeks) %>%
  summarize(times = n()) %>%
  ggplot(aes(x = weeks, y = times)) +
  geom_point() +
  geom_line() +
  ggtitle('Number of trips per week over 2013')
```
(c)
```{r echo=TRUE}
planes %>%
  select(tailnum, year) %>%
  arrange(year) %>%
  head(n=1)
```
```{r echo=TRUE}
planes %>%
  summarize(plane.num = n_distinct(tailnum))
```
(d)
```{r echo=TRUE}
weather %>%
  filter(month == 7) %>%
  ggplot(aes(x = temp)) +
  geom_histogram() +
  ggtitle('Distribution of temperature in July 2013')
```
```{r echo=TRUE, warning=FALSE, message=FALSE}
library(ggbreak)
```
```{r echo=TRUE}
weather %>%
  group_by(month) %>%
  ggplot(aes(x = month, y = wind_speed)) +
  geom_boxplot() +
  xlab('') +
  theme(axis.ticks.x = element_line(NA),
        axis.text.x = element_blank()) +
  scale_y_break(c(50, 1040))
```
```{r echo=TRUE}
quan <- quantile(weather$wind_speed, na.rm=TRUE)
iqr <- quan[4] - quan[2]

weather %>%
  filter(wind_speed >= quan[4] + 1.5 * iqr) %>%
  arrange(desc(wind_speed))
```
Important outlier: 1048.361
```{r echo=TRUE}
weather %>%
  filter(dewp != is.na(dewp) & humid != is.na(humid)) %>%
  ggplot(aes(dewp, humid)) +
  geom_point(size = 1) +
  geom_smooth() +
  ggtitle('Relationship between dewp and humid')
```
```{r echo=TRUE}
weather %>%
  summarize(cor(dewp, humid, use = 'complete.obs'))
```
Answer: As the dewp increases, so does the humid. - a positive relationship

```{r echo=TRUE}
weather %>%
  ggplot(aes(precip, visib)) +
  geom_point(size = 1) +
  geom_smooth() +
  ggtitle('Relationship between precip and visib')
```
```{r echo=TRUE}
weather %>%
  summarize(cor(precip, visib, use = 'complete.obs'))
```
Answer: Although not apparent, visib tends to decrease as the precip increases. - a weak, negative relationship

# 3
```{r echo=TRUE}
?diamonds
```
(a)
```{r echo=TRUE}
hist(x = diamonds$carat, col = 'steelblue', xlab = 'Carat')
```
(b)
```{r echo=TRUE}
qplot(x = depth, data = diamonds)
```
(c)
```{r echo=TRUE}
qplot(x = cut, y = price, data = diamonds, geom = 'violin', fill = cut)
```

# 4
```{r echo=TRUE, warning=FALSE, message=FALSE}
library(MASS)
library(tidyverse)
```
```{r echo=TRUE}
as_tibble(Cars93)
```
(a)
```{r echo=TRUE}
ggplot(data = Cars93, aes(x = MPG.highway, y = Price)) +
  geom_point(aes(size = Horsepower))
```
(b)
```{r echo=TRUE}
ggplot(data = Cars93, aes(x = MPG.highway, y = Price)) +
  geom_point(aes(size = Horsepower, color = Origin))
```
(c)
```{r echo=TRUE}
ggplot(data = Cars93, aes(x = MPG.highway, y = Price)) +
  geom_point(aes(size = Horsepower, color = Origin)) +
  stat_smooth()
```
(d)
```{r echo=TRUE}
ggplot(data = Cars93, aes(x = MPG.highway, y = Price)) +
  geom_point(aes(size = Horsepower, color = Origin)) +
  facet_grid(facets = Cars93$Origin)
```